---
title: "Practice Week 1 and 3"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Lists and data frames allow us to store and therefore analyse more complex forms of data.

In a list we can store objects of different types, e.g. numbers, text, vectors, matrices, etc and they can have different lengths or dimensions.

A data frame is more like a matrix, except that each column can have a different type.

```{r}

j <- list(name = 'Joe', salary = 55000, union = T)
j

#access the elements of list in 3 different ways
j$salary
j[['salary']]
j[[2]]

#add to a list
j$sales <- c(10400, 12300, 13700) 
j

#add via a vector index
j[[5]] <- 'Manager' 

#delete it by making it null
j[[5]] <- NULL

#as list is a vector, these are applicable
length(j)
length(j$sales) 


#accessing a list element
names(j)

#The function unlist will convert a list into a vector using the mode of the lowest common denominator (here, that is the character mode).
unlist(j) 

#By default R will give names to an unlisted object taken from the tags. We can remove them via the unname command or giving an extra argument:
unlist(j, use.names = FALSE)

#The family of apply functions (which we first met for matrices) has a version for lists known as lapply, e.g.
lapply(list(1:3, 25:29), median)

#This applies the function median to each component of the given list and returns the answers in a list sapply (for simplified apply) produces the exact same output as a vector:
sapply(list(1:3, 25:29), median)

#We can have a list withing a list
list(a = 1, b = 2, c = list(d = 3, e = 4))


#----Text concordance------#
firstpar <- "It was a bright cold day in April and the clocks were striking thirteen Winston Smith his chin nuzzled into his breast in an effort to escape the vile wind slipped quickly through the glass doors of Victory Mansions though not quickly enough to prevent a swirl of gritty dust from entering along with him"
#Suppose we want to create a function which lists the different words used and their position in the text. For example, his is used twice in positions 17 and 21.

findwords <- function(tf) {
  txt <- unlist(strsplit(tf, ' ')) # Read in the words from the text and separate into a vector
  wl <- list() # Create a list to store the words and their positions
  for(i in 1:length(txt)) { # Loop through each word
    wrd <- txt[i]  # Get the current word
    wl[[wrd]] <- c(wl[[wrd]], i) # Add its position to the list with the appropriate tag
    }
  return(wl) # Return the answer as a list
}

head(findwords(firstpar), 4)

##count words in aplhabetical order

alphawl <- function(wrdlst) {
  nms <- names(wrdlst) # Find the tags of the list
  sn <- sort(nms) # Sort them alphabetically
  return(wrdlst[sn]) # Return them
  }
alphawl(findwords(firstpar))

##sort words frequency wise

freqwl <- function(wrdlst) {
  freqs <- sapply(wrdlst, length) # Find the freq of each word
  return(wrdlst[order(freqs)]) # Return them in order
}
freqwl(findwords(firstpar)) 
```


A data frame is just like a matrix, except that each column can have a different mode.

```{r}

##-------DATA FRAMES----------#

kids <- c('Jack', 'Jill')
ages <- c(12, 10)
d <- data.frame(kids, ages) 
d

#R treats data frames as a list
mode(d)

#so we can access it the same way
d[[2]]
d$kids
d$kids[1]

#String function will give us a detailed view
str(d)


#Whilst R treats a data frame like a list, it also allows us to extract and filter in the same way as if it were a matrix.

#using the women dataset
data(women)
str(women)

women[1:2,]
women[women$height > 70,]

#dealing with NA values! Say hello to subset function again!
subset(women, weight < 130)

#turning values into NA
women[1,1] <- women[13,2] <- NA

#completely removing NA values
women[complete.cases(women),]

#adding values to column 'letters[1:15]' becomes the name of column
women2 <- cbind(women, letters[1:15])
women2

#adding values to a row
women3 <- rbind(women, c(73, 166))
women3

#adds new columns
women4 <- cbind(women, women$height * 2.54, women$weight * 0.45)
women4

#VERY HANDY COMMAND. RENAMES THE COLUMNS
names(women4) <- c('heightin', 'weightlbs', 'heightcm', 'weightkg')
women4


#mergin 2 dataframes
d1 <- data.frame(kids = c('Jack', 'Jill', 'Jillian', 'John'),
  county = c('Dublin', 'Cork', 'Donegal', 'Kerry'))
d2 <- data.frame(ages = c(10, 7, 12), 
  kids = c('Jill', 'Lillian', 'Jack'))


merge(d1, d2) #only the common names get merged
#The merge function has extra arguments by.x and by.y to tell R which variables over which to merge.


#APPLYING FUNCTIONS TO A DATASET

#applying sort function
lapply(d,sort)

#getting it back to a dataframe
data.frame(lapply(d, sort)) 



##------LOGISTIC REGRESSION--------------#

#Logistic regression is a statistical method used to predict whether a binary variable is 0 or 1 based on other explanatory variables.

#We will estimate them using the R function glm with the argument family = 'binomial'.


#The birthwt data set comes with the R package MASS. We can load it in with:

library(MASS)
help(birthwt)
str(birthwt)

logfun <- function(x) {
    glm(birthwt2$low ~ x, family = binomial)$coef
}
sapply(birthwt2[,-1], logfun)

mod <- glm(birthwt2$low ~ birthwt2$age, family = binomial)$coef
plot(jitter(birthwt2$age), birthwt2$low, col = "red")
points(birthwt2$age, 1 / (1 + exp(-(mod[1] + mod[2] * birthwt2$age))), 
  pch = 2, col = "green")
```