---
title: "Assignment1"
output: pdf_document
author: "Varun Garg 18205136"
---



## Loading the CSV

```{r}
data= read.csv("EurostatCrime2015.csv", header = TRUE)

data
```

#Size and structure of the data frame

```{r}
typeof(data)
nrow(data)
ncol(data)
```


#Add a new column called Sex.crime which contains the sum of all the crimes that have a sexual component: Rape, Sexual.assault and Sexual.violence
```{r}
data$"Sexual.Crime"  <- rowSums(cbind(data$"Sexual.assault", data$"Sexual.violence", data$"Rape"), na.rm = T )
data
```


```{r}
data = data[,!(names(data) %in% c("Rape","Sexual.violence", "Sexual.assault"))]
data
typeof(data)
nrow(data)
ncol(data)
```

```{r}

areComplete <- data[!complete.cases(data),]
areComplete[,"X", drop =FALSE]
```

```{r}
mydata = data[complete.cases(data), ]

#mydata[,"X", drop =FALSE]
#mydata

#mydata$max<-apply(mydata, 1, max)
#mydata$max
#mydata[which.max(mydata$Ireland),]
ireland <- mydata[15,]
ireland
#display <- ireland[which.max(ireland)]

suppressWarnings(colnames(ireland)[apply(ireland,1,which.max)])
suppressWarnings(colnames(ireland)[apply(ireland,1,which.min)])
colnames(ireland)[t(apply(ireland,1,sort)[1:3,])]


#dim(mydata)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
