---
title: "Week7"
output: html_document
---


We have met most simple arithmetic operations in R: +, -, etc.
R has many more (all vectorised):
exp(): exponential function
log(): natural logarithm (see also log10 and log2)
sqrt(): square root
abs(): absolute value
sin(), cos(), tan(), asin(), acos(),…: trigonometry
min(), max(), pmin(), pmax(), which.min(), which.max(): minima and maxima
sum(), prod(), cumsum(), cumprod(): sums and products
round(), floor(), ceiling(), signif(): methods for rounding numbers
```{r}
# MIN AND MAX
z <- matrix(c(1, 5, 6, 2, 3, 2), nrow = 3, ncol = 2)
z
min(z)

z1 <- z[, 1]
z2 <- z[, 2]
min(z1, z2) 

pmin(z1, z2) #pmin and pmax find parallel minima/maxima across multiple vectors or matrices.
?pmin


##DERIVATIVE AND ITEGRATION!!! DIFFERENTATION AND INTEGRATION

D(expression(exp(x^2)), 'x') #x^2 -> 2x
integrate(function(x) x^2, 0, 1)




```



R contains a large range of functions for evaluating probability distributions. These all follow a set pattern:
d for density (or pdf)
p for cumulative density (or cdf)
q for quantiles
r for random number generation
Distributions include: norm (Normal), binom (Binomial), chisq (Chi-squared), f (F-distribution), pois (Poisson), t, etc, etc
Example: plotting the pdf of a normal distribution:

```{r}
x <- seq(-4, 4, length = 100)
plot(x, dnorm(x), type = 'l')



##Sorting


x <- c(5, 13, 12, 5)
sort(x) #sort takes a vector and puts it in order.

order(x) #order gives the order of the vector x (i.e. the first element is the smallest, etc).

rank(x) #rank gives the rank order of the vector (so the 1st and 4th elements are the smallest and are given rank (1 + 2) / 2.

#Both sort and order have an extra argument decreasing for reversing the operation.

d <- data.frame(kids = c('Jack', 'Jill', 'Billy'), ages = c(12, 10, 13))
d[order(d$kids,decreasing = TRUE),  ] #name descedning order
d[order(d$kids),  ] 
d[order(d$ages),]
d[order(d$ages, decreasing = TRUE), ] #age decreasing

?order
```

Set operations
If you have a set of items stored in a vector, R can compare them using set operations:
union(x, y): Creates the union set of x and y
intersect(x, y): Creates the intersection of x and y
setdiff(x, y): Creates the set of all objects in x that are not in y
setequal(x, y): Determines if x and y are equal
c %in% y: Tests whether c is an element of y

```{r}
##LINEAR ALGEBRA!!

#There are other useful linear algebra functions:
#solve(): finds a matrix equation solution or matrix inverse
#t(): matrix transpose
#chol(), qr(), eigen(): Cholesky, QR and eigen decomposition
#det(): Determinant
#diag(): Find or set the matrix diagonal
#sweep(): Removes summary statistics from matrices.

#solve(): solve is useful for finding solutions to simultaneous equations:

A <- matrix(c(1, -1, 1, 1), 2, 2) 
b <- c(2, 4)
x <- solve(A, b)
x

#diag(): diag works to obtain the diagonal from a matrix, to set it, or to create a diagonal matrix:

M <- matrix(1:9, 3, 3)
diag(M) #will return diagonals


diag(M) <- 7 #diagonals will become 7
M

M <- diag(c(3, 1)) #set diagonals to this value
M 

#sweet() - sweep is a neat function to remove a summary statistic from a numeric data frame/matrix:

?sweep

head(swiss, 3) # Swiss Fertility and Socioeconomic Indicators (1888) Data

# centre the variables around their mean
mean.swiss <- apply(swiss, 2, mean) # calculate the column means. Same as colMeans(swiss)
swiss.mean0 <- sweep(swiss, 2, mean.swiss)
head(swiss.mean0, 3)

#Correlation

#Correlation measures the degree of relatedness between two (or more) vectors. The Pearson correlation is defined as:
cor(swiss[,1:3])

#t-test
#Usual set-up: 2 vectors (can be different lengths) where you want to test whether or not they have the same mean.
t.test(extra ~ group, data = sleep, paired = TRUE)

#Linear regression
mod <- lm(Fertility ~ ., data = swiss)
summary(mod)
```



Heads and Tails
What is the probability of getting at least 4 heads from 5 tosses of a coin?
We could do this algebraically using the probability mass function of the Binomial distribution. (1 = head, 0 = tail).

```{r}
#Binomial distribution
1 - pbinom(3, 5, 0.5) # Pr(X >= 4) = 1 - Pr(X <= 3)

#Alternatively we could simulate a large number of such coin tosses and find how often at least 4 heads occur:
large <- 100000
x <- rbinom(large, 5, 0.5)
mean(x >= 4)

#Remember x >= 4 is a logical variable so finding the mean is equivalent to finding the proportion of times such an event occurs.

#MAXIMA OF TWO N(0,1)
#Suppose we want to find the mean of the maximum of two standard normal random variables.
nreps <- 100000
sum <- 0
for(i in 1:nreps) {
  xy <- rnorm(2)
  sum <- sum + max(xy)
  } 
print(sum / nreps)

#without loops
x <- rnorm(nreps)
y <- rnorm(nreps)
maxxy <- pmax(x, y)
mean(maxxy) 

#SNAP!!! (sample)
#Suppose you play a game of snap with two players who each have a shuffled pack of cards. You both turn over cards until you reach an identical pair.

cards <- 1:52
p1hand <- sample(cards)
p2hand <- sample(cards)
snaptime <- which((p1hand - p2hand == 0) == TRUE)[1]
snaptime
#Here it took 24 draws before snap was called

```
