---
title: "Week8"
output: html_document
---

Everything we use in R (functions, vectors, data frames, etc) is an object.
There are several important features of OOP that R promotes:
Encapsulation - data items are packaged into one class instance and makes keeping track of everything easier.
Polymorphism - the same function will have different actions on different classes.
Inheritance - objects of one class can inherit characteristics of another class.
R mainly uses two types of classes: S3 and S4. In this lecture we'll study them in detail and go through the differences

```{r}

#An S3 class consists of a list with a class name attribute (e.g. lm) and dispatch capability.
#The dispatch capability means that you can make use of generic functions (e.g. print or summary).


#S3 generic functions
#We have met a number of functions that work on vectors, lists, matrices, etc: print, plot, summary.

#These are known as generic functions, in that they act differently for different types of objects given to them.
y <- list(tag1 = 2, tag2 = c(2, 3), tag3 = 1:10)
print(y)

mod <- lm(Fertility ~ ., data = swiss)
print(mod) #will act differently

class(y) #list
class(mod) #lm

unclass(mod)

j <- list(name = 'Joe', salary = 55000, union = TRUE)
class(j) <- 'employee'
j 

print.employee <- function(wkr) {
  cat(wkr$name, '\n')
  cat('salary', wkr$salary, '\n')
  cat('union member', wkr$union, '\n')
}
j

x <-  list(name = 'Jj', salary = 5100, union = TRUE)
class(x) <- 'employee'
x

k <- list(name = 'Kate', salary = 68000, union = F, hrs_this_month = 2)
class(k) <- c('hrly_employee', 'employee') #inheriting employee
k

summary(k)


```


Problems with S3
S3 classes do not provide much safety if we make a mistake.
For our example on employees (slide 10), we might:
Forget to enter a field.
Mis-spell one of the fields.
Accidentally give a different object the class employee.
An object of S3 class will not complain when such things happen, but an object of S4 class will not allow such mistakes.
S4 classes are much more formal when we are creating objects, defining the class, etc.

```{r}
#The setClass function will create an S4 class, whilst new will create S4 objects:

#make a class first
setClass('employee', 
  representation(name = 'character', salary = 'numeric', union = 'logical'))
#assign an object to it
joe <- new('employee', name = 'Joe', salary = 55000, union = TRUE)
joe

#S4 objects contain values in slots rather than tags.

#Slots are accessed with an @ sign or the slot function:

joe@salary
slot(joe, 'union')

#We can update objects in the same way as a list:
joe@salary <- 65000 

#… but it won't allow us to add new elements:

joe$salary <- 55000


#The setMethod function creates new S4 generic functions:

setMethod('show', 'employee', 
  function(object) {
  inorout <- ifelse(object@union, 'is', 'is not')
  cat(object@name, 'has a salary of', object@salary, 'and', 
    inorout, 'in the union \n')
  }
)
joe

```